---
title: "Day 3 Pset"
author: "Cecilia Sui"
date: "1/14/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Day 3 Outline:
1. Packages and Help Pages 
2. Data Wrangling Part 2
3. Tips on Using R Markdown


# dplyr Basics 
We will use the resources dataset to do some exercises with dplyr
```{r include=FALSE}
library(dplyr)
library(ggplot2)
df <- read.csv("/Users/river/Documents/GitHub/RCamp2023/Day5/Pset/resources.csv")
colnames(df)
```

1. Select the first 10 rows and last 3 rows. Output the rows together as the
   same dataframe.
```{r}
dim(df)
test <- df %>% slice(c(1:10, 4528:4530))
test
```


2. Select all rows where Year is 1951 and regime is non-negative.
```{r, eval=FALSE}
df %>% filter(year == 1971, regime >= 0)
```

3. Order the rows by Year in descending order. Break ties using the column oil in ascending order.
```{r, eval=FALSE}
df %>% arrange(desc(year), oil)
```

4. Select the 100th to 110th rows, and find the row with the highest value for
   metal.
```{r}
df %>% slice(100:110) %>%
    slice_max(metal)
```


5. Randomly select 10 rows with replacement, and standardize the columns oil and metal.
```{r}
test <- df %>%
  na.omit() %>% # This way we are only dealing with complete cases
  slice_sample(n = 10, replace = TRUE) %>%
  mutate(
    oil_sd = (oil - mean(oil)) / sd(oil),
    metal_sd = (metal - mean(metal)) / sd(metal)
  )
```


6. Select columns: cty_name, year, regime, oil, illit, and life and store them
   into a new dataframe.
```{r}
new_df <- df %>%
  select(cty_name, year, regime, oil, illit, life)
```

7. Add a new column named new_oil using where the values are the product of standardized oil and regime type.
```{r}
new_df <- new_df %>%
  mutate(new_oil = ((oil - mean(oil, na.rm = TRUE)) / sd(oil, na.rm = TRUE)) * ((regime - mean(regime, na.rm = TRUE)) / sd(regime, na.rm = TRUE)))
```

8. Calculate the **average** oil level for **each year** from 1966 to 1997. 
```{r}
vec <- c()
j <- 1
for (i in 1966:1997) {
  vec[j] <- mean(new_df$oil[new_df$year == i], na.rm = TRUE)
  j <- j + 1
}
vec

new_df %>%
  group_by(year) %>%
  summarise(oil = mean(oil, na.rm = TRUE))
```

9. Calculate the **median** metal value for **each** regime type from -10 to +10.
```{r}
df %>%
  group_by(regime) %>%
  summarise(metal = median(metal, na.rm = TRUE))
```
