---
title: "Probability Solutions"
modified: 
output: 
  pdf_document: default
---

# Probability Distributions

1. Suppose the height of males at a certain school is normally distributed with
   a mean of $\mu=168$ cm and a standard deviation of $\sigma = 9$ cm.
   Approximately what percentage of males at this school are **taller** than 175
   cm?
```{r}
pnorm(175, mean = 168, sd = 9, lower.tail = FALSE)
```

2. Suppose the weight of a certain species of dogs is normally distributed with
   a mean of $\mu=33$ lbs and a standard deviation of $\sigma = 5.5$ lbs.
   Approximately what percentage of this species of otters weight less than 28
   lbs?
```{r}
pnorm(28, mean = 33, sd = 5.5, lower.tail = TRUE)
```

3. Suppose the height of flowers in a certain region is normally distributed
   with a mean of $\mu=24$ cm and a standard deviation of $\sigma= 5$ cm.
   Approximately what percentage of flowers in this region are **between 18 and
   25 cm** tall?
```{r}
pnorm(25, mean = 24, sd = 5) - pnorm(18, mean = 24, sd = 5)
```

4. Find the Z-score of the 99th quantile of the standard normal distribution.
```{r}
qnorm(0.99)
```

5. Find the Z-score of the 95th quantile of the standard normal distribution.
```{r}
qnorm(0.95)
```

6. Generate a vector of 5 normally distributed random variables with mean = 10
   and sd = 2
```{r}
set.seed(12345)
rnorm(5, mean = 10, sd = 2)
```
7. Generate a vector of 100 normally distributed random variables with mean = 50
   and sd = 5
```{r}
rnorm(1000, mean = 50, sd = 5)
```

8. Generate a vector of 1000 normally distributed random variables with mean =
   50 and sd = 25
```{r}
rnorm(1000, mean = 50, sd = 25)
```
9. Generate two histograms using the **hist()** function to view these two
   distributions.
```{r}
hist(rnorm(100, mean = 50, sd = 25))
hist(rnorm(1000, mean = 50, sd = 25))
```
